Imbalanced class distribution is a common issue in many classification tasks. For example, when classifying whether a patient has cancer or not, it is not unusual that the training data contains highly imbalanced positive/negative examples which however reflects the true class distribution in a general population. In predictive maintenance applications, the same issue exists: the imbalance of failure events to normal operation events. This issue is due to following two major reasons. First, the failure events usually rarely occurs compared to normal operation state for an in-service asset. Second, there are too few failure events. The business cannot afford to let the asset run-to-failure, as it is at the cost of equipment damage and equipment down time.

The SMOTE module is created based on algorithm "SMOTE: synthetic minority over-sampling technique" [2]. It is used to increase the size of the minority examples in a data set by synthesizing new examples with minority class. SMOTE module has two parameters: "SMOTE percentage" and "Number of nearest neighbors". The parameter "SMOTE percentage" should be in multiples of hundreds (100,200,300,400,…). This is fraction of new minority examples that gets added. For examples, we double our minority class by setting the value to 100, we triple the size of minority class by setting the value to 200, etc. The parameter "Number of nearest neighbors" is used to generate new examples from minority class. Each generated example is an average of the original example and its nearest neighbors from the same class.